Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex
Related work section

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstracting
    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
    describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited controls,
    we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same
    definitional interpreter with pushdown control flows. In this case, nondeterminism can be
    easily handled for comprehension.
\end{itemize}

@@

\textbf{Defunctionalization and Refunctionalization.}
Defunctionalization and refunctionalization build connections between abstract
machines and interpreters.
\citeauthor{reynolds1972definitional} first shows refunctionalization as a program transformation technique that 
can be used to transform higher-order functions to first-order functions 
when sepcifing a programming language by interpreter \cite{reynolds1972definitional}.
Refunctionalization is the reversed transformation.

Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization to 
many different abstract machines and semantics, including the CEK machines, the CLS machines, 
the SECD machines, and etc
\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
And other applications are also found to be applicable, such as Dyck words recognizer and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.

For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows 
a reduction semantics is a structural operational semantics in continuation style, 
where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
Our work further shows for nondeterministic languages, there is another layer of continuations
that controls the nondeterministic choices. After identifing and explicitly exposing (section \ref{linear}) that layer
of nondeterministic continuations, the whole program can be then refunctionalized to extended continuation-passing style.

>>>>

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstracting
    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
    describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited controls,
    we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same
    definitional interpreter with pushdown control flows. In this case, nondeterminism can be
    easily handled via @for@ comprehension.
\end{itemize}

@@

\textbf{Defunctionalization and Refunctionalization.}
Defunctionalization and refunctionalization build connections between abstract
machines and interpreters.
\citeauthor{reynolds1972definitional} first shows defunctionalization as a program transformation technique that
can be used to transform higher-order functions to first-order functions
when specifying a programming language by an interpreter \cite{reynolds1972definitional}.
Refunctionalization is the reverse of this transformation.

Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization to
many different abstract machines and semantics, including the CEK machines, the CLS machines,
the SECD machines, etc.
\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}
Other applications are also found to be applicable, such as Dyck word recognizer, and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.

For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows
a reduction semantics is a structural operational semantics in continuation style,
where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
Our work further shows that for nondeterministic languages, there is another layer of continuations
that controls the nondeterministic choices. After identifying and explicitly exposing (Section~\ref{linear}) that layer
of nondeterministic continuations, the whole program can then be refunctionalized to extended continuation-passing style.
